---
title: "DK Forest LiDAR - By Pixel Projections of Forest Quality"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(raster)
library(tidyverse)
library(leaflet)
library(caret)
library(gbm)
library(sf)
```


[Fast forward to the bottom](#aarhus-region) of this document for the map. 

## Boosted Regression Tree Model
```{r }
# Load data
load("D:/Jakob/dk_forest_lidar_2021/data/pixel_sample_10k.Rda")

# Reshape data 
combined_sample <- combined_sample %>% map(function(x) dplyr::select(x, -forest_class)) %>%
  reduce(full_join, by = "sample_id") %>% full_join(combined_sample[[1]][,1:2], 
                                                    ., by = "sample_id")
# Filter out masked data and na values   
combined_sample <- combined_sample %>% 
  filter(!is.na(inland_water_mask)) %>%
  filter(!is.na(sea_mask)) %>%
  select(-contains("mask")) %>%
  mutate(forest_class = factor(forest_class)) %>%
  na.omit()

# Remove - from variable names
names(combined_sample) <- gsub("-", ".", names(combined_sample))

# Slit dataset into training and control
set.seed(32423)
index_training <- createDataPartition(
  y = combined_sample$forest_class,
  p = 0.8,
  list = F
)
train_data <- combined_sample[index_training, -1]
test_data <- combined_sample[-index_training, -1]

# Load model 
load("D:/Jakob/dk_forest_lidar_2021/data/final_gbm_model_pixel.Rda")
```

The model uses the EcoDes-DK15 dataset to predict forest quality. Annotations of forest quality were provided by polygons from various Danish agencies:

| forest quality | annotation source |
| --- | --- |
| high | ยง15 forest polygons |
| high | ยง25 forest polygons |
| high | private old growth |
| low | ikke ยง25 forests polygons |
| low | NST plantation polygons |

The polygons were bagged together in group by forest quality class. We then obtained a simple random sample of 10k pixels across Denmark for each group. 

``` {r }
# Load shapefiles
high_quality <- list.files("../data/response/high_quality_forests/", 
                           ".shp", 
                           recursive = T,
                           full.names = T) %>%
  map(function(x) {
    sf_object <- read_sf(x)
    sf_object <- sf_object %>%
      select(!everything()) %>%
      mutate(forest_id = paste0(gsub(".*/(.*)\\.shp", "\\1", x),
                                "_", 1:n()),
             forest_class = "high")
    }) %>%
  bind_rows() 
low_quality <- list.files("../data/response/low_quality_forests/", 
                          "shp$", 
                          recursive = T,
                          full.names = T) %>%
  map(function(x) {
    sf_object <- read_sf(x)
    sf_object <- sf_object %>%
      select(!everything()) %>%
      mutate(forest_id = paste0(gsub(".*/(.*)\\.shp", "\\1", x),
                                "_", 1:n()),
             forest_class = "low")
  }) %>%
  bind_rows() 
combined <- rbind(high_quality, low_quality) %>% st_transform(4326)

# # Make leaflet
# m <- leaflet(combined) %>% setView(lng = 10.208681, lat = 56.156397, zoom = 10) %>% 
#   addProviderTiles(providers$Esri.WorldImagery) %>%
#     addPolygons(
#     color = ~forest_class,
#     fillOpacity = 0.5
#   )

```

The resulting 20k pixel samples from the EcoDes-DK15
dataset were then used to train three types of binary classifier models (random forests, gradient boosted regression trees and a simple multilayer perceptron nural network). Prior hyperparameter tuning,
the three model types performed similary and we chose to progress with the boosted regression tree model.

We used 80% of the pixels for training and kept 20% aside for validation (simple random split). Addtionally, the modeltraining was carried out using 10-fold repeat crossvalidation with the 80% training data. 

The overall model performance is good ~ 82% accuracy. See details below:

``` {r }
# Validate on test set
test_preds <- predict(gbm_fit, newdata = test_data)
confusionMatrix(data = test_preds, test_data$forest_class)
```

Here are the variables that make up the 20 most important predictors:
``` {r}
# Check variable importance
varImp(gbm_fit)
```

## Aarhus Region {#aarhus-region}
Here is a projection of the model results for the Aarhus region:

``` {r warning = FALSE}
forest_qual <- raster("D:/Jakob/dk_forest_lidar_2021/data/projections/by_pixel/forest_quality_by_pixel.vrt")
pal <- colorFactor(c("blue", "red"), domain = c("High" = 1,
                                                "Low" = 2),
  na.color = "transparent")

leaflet() %>% 
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addRasterImage(forest_qual, colors = pal, opacity = 1,
                 method = "ngb", group = "forest_qual") %>%
  addLegend(pal = pal, values = c(1,2),
            labFormat = labelFormat(transform = function(x){
              c("High", "Low")[x]
            }),
    title = "Projected Forest Quality by Pixel") %>%
    addLayersControl(
    overlayGroups = c("forest_qual"),
    options = layersControlOptions(collapsed = FALSE)
  )
```

