---
title: "Visualising large rasters with Leaflet - a quick guide"
author: "Jakob J. Assmann"
contact: j.assmann@bio.au.dk
date: "10/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stars)
library(sf)
library(terra)
library(leaflet)
```

### Overview

This is a quick guide on how to visualise a cloud optimised GeoTiff using a Leaflet.

Workflow overview:

1. [Generate a cloud otimised geotiff.](#generate-a-cloud-otimised-geotiff)
2. [Host the raster on the cloud.](#host-the-raster-on-the-cloud)
3. [Add the raster to a Leaflet.](#add-the-raster-to-a-leaflet) 

There are various ways of how each step can be achieved. The guide below sums up what worked best for me. 

I got my inital inspiration from Sean Rennie's [post on medium](https://sean-rennie.medium.com/cogs-in-production-e9a42c7f54e4). 

### 1) Generate a cloud otimised geotiff

Hosting ridiculously large raster has become a lot easier since the introduciton 
of cloud optimised GeoTiffs (often abbreviated to COGs). 
To learn more about COGs see [cogeo.org](https://www.cogeo.org/). 
A relatively mature driver is already implemented in [GDAL](https://gdal.org/drivers/raster/cog.html), 
which we will use here to generate a COG from our raster. 

The advantage of COGs is that they directly integrate a pyramid of overview tiles into the 
GeoTiff file (no sidecar files needed) and that they allow for HTTP GET request to access 
only those tiles that are needed for displaying / working with the raster at a certain scale. 
Overall, this allows for fast and efficient visualisation of the raster
while not obscuring direct access to the raw raster values. 

#### Generating the COG

There are multiple routes to generate a COG with GDAL. File compression and tiling can be
added to a GeoTiff in a stepwise manner using `gdal_warp` and `gdal_addo`, but 
I found that the following one-step conversion with `gdal_warp` worked best for me:

```
gdal_warp -of COG -co RESAMPLING=NEAREST -co TILING_SCHEME=GoogleMapsCompatible 
  -co COMPRESS=DEFLATE -co NUM_THREADS=46 source_file.tif out_file.tif
```

Here, I am using a nearest neighbour resampling algorithm (option 
`RESAMPLING`) and 46 parallel threads (option `NUM_THREADS`). Just adapt these 
parameters according to the data type / computing resources available to you.

The `DEFLATE` compression provided great results for me and seems to be handled
efficiently by the georaster layer for Leaflet. While some other guides suggest
adding tiles manually, I found that using the `GoogleMapsCompatible` tiling scheme
provided the best results and fastest response, the GDAL driver also automatically 
seems to project the raster into EPSG 3857 (web mercartor) when using this option. 

More detail on the driver options can be found in the GDAL documentation linked 
above.

If the raster needs to be generated from within R, I suggest using `gdal_utils` 
from the `sf` package. I often use the gdal binaries shipped with `sf`,
as accessing them is easy and I generally have the package loaded when 
working with spatial data in R. 

The sf::gdal_utils call for the above GDAL command would look as follows:

```{r eval = FALSE}
gdal_utils("warp",
           source = "source_file.tif",
           destination = "out_file_cog.tif",
           options = c(
             "-of", "COG",
             "-co", "RESAMPLING=NEAREST",
             "-co", "TILING_SCHEME=GoogleMapsCompatible",
             "-co", "COMPRESS=DEFLATE",
             "-co", "NUM_THREADS=46"
           ))
```

#### Let's try it out! 

First we generate a random binary raster (values 1 or 2) that spans Denmark in the projected CRS "UTM 32N":

```{r, eval = FALSE}
library(terra)
library(sf)
library(rnaturalearth)
# Retrieve a polygon of Denmark via rnaturalearth and project to UTM 32N
denmark <- ne_countries(scale = "medium", 
                        country = "denmark",
                        returnclass = "sf") %>%
  st_transform(25832)
# Retrieve extent, add a buffer and adjust to allow for clean divison by 100 m 
extent_dk <- (st_bbox(denmark) - (st_bbox(denmark) %% 100) + 100) %>%
  .[c("xmin", "xmax", "ymin", "ymax")] %>%
  as.numeric() %>%
  ext()
# Generate a random 100 m res raster (1s and 2s), mask out everything outside DK
random_rast_dk <- rast(nrows = (extent_dk[4] - extent_dk[3]) / 100,
                       ncol = (extent_dk[2] - extent_dk[1]) / 100,
                       extent = extent_dk,
                       crs = "EPSG:25832")
values(random_rast_dk) <- rbinom(ncell(random_rast_dk), 1, 0.75) + 1
random_rast_dk <- mask(random_rast_dk, vect(denmark))
# Save as an ordinary tiff file
writeRaster(random_rast_dk, "random_rast_dk.tif")
```

Now we convert the raster (3.1 Mb) to a compressed COG in EPSG 3857 using `gdal_utils()`:

```{r eval = FALSE}
gdal_utils("warp",
           source = "random_rast_dk.tif",
           destination = "random_rast_dk_cog_epsg_3857.tif",
           options = c(
             "-of", "COG",
             "-co", "RESAMPLING=NEAREST",
             "-co", "TILING_SCHEME=GoogleMapsCompatible",
             "-co", "COMPRESS=DEFLATE",
             "-co", "NUM_THREADS=6"
           ))
```

Note how the resulting COG has a reduced file size (2.5 MB). For bigger rasters
the compression can be even more notable (I had ~6GB rasters reduced to ~40 MB). 
Next we look at how to host the COG properly on the cloud. 

### 2) Host the raster on the cloud

This step was the one that took me the longest to figure out. I only dabble in 
webhosting and it was a steep learning curve. 

The biggest issue
was that the georaster for leaflet plugin (see below) needs the hosting service 
to allow for Cross Origin Resource Sharing (CORS). Both AWS and Google Cloud 
block CORS by default. While this is a reasonable measure, for making the COG
to work with leaflet CORS needs to be enabled for the file. 

Here I provide a quick run through on how a COG can be hosted using an AWS S3 bucket. 
The steps required for the Google Cloud should be similar, I suggest you consult 
the Google Cloud documentation if you would like to use Google Cloud instead. 

In the following section, I assume that you have already fully [set up an AWS account with S3](https://docs.aws.amazon.com/AmazonS3/latest/userguide/setting-up-s3.html) and 
that you are familiar with the [basics of S3 file storage](https://docs.aws.amazon.com/AmazonS3/latest/userguide/GetStartedWithS3.html).

All steps can be carried out in the [AWS S3 web console](https://s3.console.aws.amazon.com/). 

1. Create a new bucket (or modify an old one with now sensitive data).
2. Go to "Permissions" and change the "Block public access" settings for the bucket so that they allow for "public access through access control lists (ACLs)". You can also set this during bucket creation. 
3. Still in the "Permissions" tab for the bucket, scroll down to "Cross-origin resource sharing (CORS)".
4. Click "Edit" and paste the following JSON into the text field:
```
[
    {
        "AllowedHeaders": [
            "*"
        ],
        "AllowedMethods": [
            "GET",
            "HEAD"
        ],
        "AllowedOrigins": [
            "*"
        ],
        "ExposeHeaders": []
    }
]
```
5. Save the changes. 

**Note: The above JSON allows for all "GET" and "HEAD" CORS requests to access all 
[header types provided by S3](https://docs.aws.amazon.com/AmazonS3/latest/API/RESTCommonRequestHeaders.html) independent of their origin ("\*"). To learn more about the AWS CORS configuration see the 
[AWS docs help page on CORS](https://docs.aws.amazon.com/AmazonS3/latest/userguide/ManageCorsUsing.html). 
It might be worthwhile reading into the documentation if you would like to 
tighten up the security of your bucket a bit more. For the purposes of hosting my GeoTiffs, these
settings were safe enough for me, while at the same time allowing the leaflet 
to work nicely.**

6. Upload the COG to the bucket (e.g., the above generated "random_rast_dk_cog_epsg_3857.tif"). 
7. In the file's "permission" tab, set the ACL permission to "Read" for the group "Everyone (public access)".

We are now ready to access the COG on S3 using Leaflet. Before you 
leave the S3 web console don't forget to copy the AWS S3 URL to the raster  
into your clipboard and store it somewhere for later use. 
(e.g. "https://dkforestlidar2022.s3.eu-central-1.amazonaws.com/random_rast_dk_cog_epsg_3857.tif")

### 3) Add the raster to a Leaflet 

The final step is to add the COG hosted on S3 to a Leaflet. The functionality of
adding a COG to a Leaflet is provided by the JS plugin [georaster-layer-for-leaflet](https://github.com/GeoTIFF/georaster-layer-for-leaflet) 
by Daniel Dufour. I highly recommend to check out the documentation and the [examples](https://geotiff.github.io/georaster-layer-for-leaflet-example/). 
The latter provide some inspiration on the variety of ways on how the plugin can be used. 
The examples are also very helpful for understanding how colour mapping and 
spectral index calculations can be realised using the plugin.

For the dk_forest_lidar project, I wrote a JS script / website that generates a 
[realtively complex Leaflet web app](https://jakobjassmann.github.io/dk_forest_lidar_2021/data_vis.html).
You can check out the [source code here](https://github.com/jakobjassmann/dk_forest_lidar_2021/blob/main/docs/data_vis.html).
The functionality of the web app exceeds what is needed for the purpose of this guide. I will keep it more simple here. 

***Note: The [leafem R package](https://github.com/r-spatial/leafem) is currently
developing support for the Geotiff API - see [Issue 29](https://github.com/r-spatial/leafem/issues/29). The current implementation is still iffy and I can't get it to work reliably with COGs. 
Therefore I am directly adding HTML and JS code to the R markdown file here instead
of using the leaflet package for R.***

We start by loading all the Java Script dependencies for leaflet, georaster and
the geo-raster-for-leavlet plugin. The code boxexs below show the code that has 
been added to the raw text of this section. Check out the [source file](https://github.com/jakobjassmann/dk_forest_lidar_2021/blob/main/docs/cog_guide.Rmd) for this 
document if you are not clear on what this means. 

```{HTML}
<!--- Leaflet CSS stylesheet (needs to be loaded before leaflet js) --->
<link rel="stylesheet" href="https://unpkg.com/leaflet@1.7.1/dist/leaflet.css" integrity="sha512-xodZBNTC5n17Xt2atTPuE1HxjVMSvLVW9ocqUKLsCC5CXdbqCmblAshOMAS6/keqq/sMZMZ19scR4PsZChSR7A==" crossorigin="" />

<!--- Leaflet js --->
<script src="https://unpkg.com/leaflet@1.7.1/dist/leaflet.js" integrity="sha512-XQoYMqMTK8LvdxXYG3nZ448hOEQiglfqkJs1NOQV44cWnUrBc8PkAOcXy20w0vlaXaVUearIOBhiXZ5V3ynxwA==" crossorigin=""></script>

<!--- Georaster js --->    
<script src="https://unpkg.com/georaster"></script>

<!--- Georaster-Layer-For-leaflet js --->      
<script src="https://unpkg.com/georaster-layer-for-leaflet/dist/georaster-layer-for-leaflet.min.js"></script>
```

<link rel="stylesheet" href="https://unpkg.com/leaflet@1.7.1/dist/leaflet.css" integrity="sha512-xodZBNTC5n17Xt2atTPuE1HxjVMSvLVW9ocqUKLsCC5CXdbqCmblAshOMAS6/keqq/sMZMZ19scR4PsZChSR7A==" crossorigin="" />

<script src="https://unpkg.com/leaflet@1.7.1/dist/leaflet.js" integrity="sha512-XQoYMqMTK8LvdxXYG3nZ448hOEQiglfqkJs1NOQV44cWnUrBc8PkAOcXy20w0vlaXaVUearIOBhiXZ5V3ynxwA==" crossorigin=""></script>

<script src="https://unpkg.com/georaster"></script>

<script src="https://unpkg.com/georaster-layer-for-leaflet/dist/georaster-layer-for-leaflet.min.js"></script>

To make the leaflet work fully in this RMarkdown file, I then load an 
empty leaflet. This adds the CSS style sheets from the leaflet package to the 
headers of the generated HTML and ensures that any leaflets we add are displayed
as expected. 

```{r, warning = FALSE}
library(leaflet)
leaflet(width = 0, height = 0)
```
Next we add a div element for the actual map. 
```{HTML}
<div id = "map" style="width: 100%; height: 480px; position: relative;"></div>
```

The map will appear here:

<div id = "map" style="width: 100%; height: 480px; position: relative;"></div>
<br>
Finally, we add a plain text JS section to generate the leaflet and load the 
raster stored on the S3 bucket:

```{HTML}
<script>
  // Load base tiles ----

  // OpenStreetMap Tiles
  var OpenStreetMap = L.tileLayer('https://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png', {
    attribution: '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors'
  });

  // Initiate leaflet ----
  // This needs to be done before loading the Georasters
  var map = L.map('map', { 
      center: [56.25, 12], // Centered on Denmark
      zoom: 7, // Worked well
      layers: [OpenStreetMap] // Default layers
  });

  // Set URL to cog
  var random_rast_dk_url = "https://dkforestlidar2022.s3.eu-central-1.amazonaws.com/random_rast_dk_cog_epsg_3857.tif"

  //  Load georaster layer and add to map
  var random_rast_dk_url = parseGeoraster(random_rast_dk_url).then(georaster => {
      var random_rast_dk = new GeoRasterLayer({
          georaster,
          opacity: 0.75,
          resolution: 128, // 128 seems to be a good compromise
          pixelValuesToColorFn: function(pixelValues){
              var pixelValue = pixelValues[0]; 
              if (pixelValue >= 1 && pixelValue <= 2){
                  var colour = (pixelValue == 1 ? "Purple": "DarkGreen"); 
                  return colour;
              } else {
                  return undefined;
              }   
          }
      });       
      random_rast_dk.addTo(map);
  });

</script>
```

<script>
  // Load base tiles ----

  // OpenStreetMap Tiles
  var OpenStreetMap = L.tileLayer('https://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png', {
    attribution: '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors'
  });

  // Initiate leaflet ----
  // This needs to be done before loading the Georasters
  var map = L.map('map', { 
      center: [56.25, 12], // Centered on Denmark
      zoom: 7, // Worked well
      layers: [OpenStreetMap] // Default layers
  });

  // Set URL to cog
  var random_rast_dk_url = "https://dkforestlidar2022.s3.eu-central-1.amazonaws.com/random_rast_dk_cog_epsg_3857.tif"

  //  Load georaster layer and add to map
  parseGeoraster(random_rast_dk_url).then(georaster => {
      var random_rast_dk = new GeoRasterLayer({
          georaster,
          opacity: 0.75,
          resolution: 128, // 128 seems to be a good compromise
          pixelValuesToColorFn: function(pixelValues){
              var pixelValue = pixelValues[0]; 
              if (pixelValue >= 1 && pixelValue <= 2){
                  var colour = (pixelValue == 1 ? "Purple": "DarkGreen"); 
                  return colour;
              } else {
                  return undefined;
              }   
          }
      });       
      random_rast_dk.addTo(map);
  });

</script>

For access to the raster (e.g., to do some calculation in a Shiny app), you can 
load the raster remotely as a proxy using stars:

```{r warnings = FALSE, messages = FALSE}
library(stars)
# Load remote raster
random_rast <- read_stars("https://dkforestlidar2022.s3.eu-central-1.amazonaws.com/random_rast_dk_cog_epsg_3857.tif",
                          proxy = T,
                          quite = T)
# Extract a point
st_sfc(st_point(c(12, 56.25)), crs = 4326) %>%
  st_transform(st_crs(random_rast)) %>%
  st_extract(random_rast, .,
             quite = T) %>%
  st_as_sf()
```

