---
title: "DK Forest LiDAR - Predictor Data Overview"
author: "Jakob J. Assmann"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output: 
    html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(raster)
library(sf)
library(tidyverse)
library(terra)
library(kableExtra)
```

## Predictor variables and selection

### EcoDes-DK15 descriptors

The core of the predictor variables is formed by the EcoDes-DK15 rasterised 
lidar descriptors (Assmann et al. 2022) generated from the 2014/15 national
airborne laser scanning campaign conducted by the Danish government.

From the 76 available EcoDes-DK15 layers (incl. auxiliary layers), 
we removed the date_stamp_xxx, point_count_xxx, point_source and 
building_proportion layers as we deemed those non-informative for the task of 
predicting forest quality. We kept the sea and water mask layers to try out
sub-setting of the training data to make sure only land pixels are included, 
but discarded the mask layers later in the analysis. 

Furthermore, we removed the following descriptors: canopy_openness, point_count, 
normalized_z_mean, heat_load_index, openness_mean, twi -
as the ecological meaning of these was conceptually redundant with other 
descriptors (vegetation_density, canopy_height, solar_radiation, 
openness_difference and ground water respectively) and initial model runs 
indicated that these variables had a low predictive power. We also removed 
the aspect variable because it was a very weak predictor. This makes sense 
conceptually as the aspect at 10 m likely has little meaning on whether a 
forest cell is of high quality or not (all cardinal directions would 
theoretically be expected to be of high quality). 

Finally, we removed all vegetation_proportion variables. These variables 
demonstrated a low predictive power by themselves. However, to capture the 
vertical variability in the lidar point cloud we calculated a foliage height
diversity variable.

The final set of used EcoDes-DK15 variables is:

- amplitude_mean
- amplitude_sd
- canopy_height
- dtm_10m
- normalized_z_sd
- openness_difference
- slope
- solar_radiation
- vegetation_density 

### Foliage height diversity

To capture the vertical variation in the forest canopy we calcualted the
"foliage height diversity" (MacArthur and MacArthur 1961) from the EcoDes-DK15
point proportion descriptors We followed the height bins used by 
Wilson (1974): 0 m – 1.5 m, 1.5 m – 9 m, and >9 m.

- foliage_height_diversity

### Tree type predictor

As we expected that most common tree type (broadleaf vs. coniferous) would play 
an important role in determining if and why a forest is of high or low quality, 
we included the tree type projections generated by Bjerreskov et al. (2021). 

The authors used a multi-temporal Sentinel 1/2 data fusion (SAR and optical) 
approach to assign forest types in a binary classification (broadleaf vs. coniferous).

As both types are mutually exclusive we discarded the "is confierous" variable after
one-hot encoding of the source data. The source data is currently not publcily 
avialable, but was kindly shared with us by Thomas Nord-Larsen (senior author 
on Bjerreskov et al. 2021).

- treetype_bjer_dec

### Soil predictors

#### Clay, sand and organic carbon content of soil

Soil type and composition are an important indicator in the key for the paragraph
25 forests. Here we used the following three predictors to account for 
differences in the soils across Denmark:

- Clay_utm32_10m
- Sand_utm32_10m
- Soc_utm32_10m

These data were obtained from the Soilgrids 2.0 dataset (Poggio et al. 2021).
The original data layers were queried using the geodata package
(Hijmans, Ghosh, and Mandel 2021) and subset to the extent of Denmark. The original 
data have a grain size of 250 m and are in a "Interrupted_Goode_Homolosine" projection.
We projected them to the EcoDes-DK grid with 10 m grain size (UTM32N) using 
nearest neighbor resampling. 

Note that the nearest neighbour resampling strategy is conservative and makes no 
assumption about the spatial distribution of the variables during the downsampling 
of the 250 m dataset. However, the downsampling may give the wrong impression that we
have used higher-resolution predictor data than we actually have. Finally, the resampling
will inevitably introduce some uncertainties where the downsampled grid and the orignal grid
not align.

As a water mask had originally been applied to this data, we had not predictor data
were a 250 m x 250 m pixel overlapped with a water body. This became a problem when
extrapolating the models to the nationwide extent, as the finer grain size of our maps
introduced more detailed shore lines. We therefore had 10 m x 10 m land pixels for which
no soil data was available. To address this problem we gap-filled the original 
250 m x 250 m soil data. All pixels that were NA with at least one neighbouring cell
that was not NA were filled with the mean of all neighbouring cells that are not NA. 
The raster was then projected to the EcoDEs-DK grid with 10 m grain size and
only used for generating the nationwide forest quality maps from the trained
models, but not for training of the models. Forest Quality predictions close to some
shores may therefore contain some error, but we are confident that this error is
very small due to the inherently high autocorrelation of the soil variables. 

#### Water availability 

To account for the wetness of the forest ground and the water availability to the plants
we use the summer near-surface ground water estimates by Koch et al. 2021.

- ns_groundwater_summer

### Focal variables

To capture the spacial context around a pixel beyond the 10 m grid, we selected 
four key predictor variables and calculated their mean and variation (sd) for 
two window sizes of 110 m and 250 m around each pixel. We selected these window 
sizes as the best candidates based on variograms generated for all variables.

We conducted a collinearity analysis on the focal variables and reduced the v
ariables in a step-wise selection process to the following final four focal 
variables included in the models:

- dtm_10m_sd_110m
- canopy_height_sd_110m
- vegetation_density_sd_110m
- ns_groundwater_summer_sd_110m

Additional documentation of the selection process can be found in the 
[focal variable selection](focal_var_selection.html) document. 

## Overview table final predictor data sources

Here is an overview table of the final predictor data sources. 

```{r echo=FALSE}
# Load data
load("../data/training_data/pixel_training.Rda")

# Filter out masked data and na values   
predictor_vars <- pixel_training_data %>% 
  st_drop_geometry() %>%
  select(-contains("mask")) %>%
  select(-contains("250")) %>%
  select(-contains("mean_110")) %>%
  select(-contains("proportion")) %>%
  select(-heat_load_index,
         -aspect,
         -openness_mean,
         -normalized_z_mean,
         -canopy_openness,
         -treetype_bjer_con,
         -twi,
         -forest_value,
         -sample_id,
         -biowide_region,
         -dereks_stratification,
         -terron_point) %>% 
  names() %>%
  sort()

# Build dataframe
tibble(Predictor = predictor_vars,
           `Source Dataset` = c("EcoDes-DK15",
                                "EcoDes-DK15",
                                "EcoDes-DK15",
                                "EcoDes-DK15",
                                "Poggio et al. 2021",
                                "EcoDes-DK15",
                                "EcoDes-DK15",
                                "EcoDes-DK15",
                                "EcoDes-DK15",
                                "Koch et al. 2021",
                                "Koch et al. 2021",
                                "EcoDes-DK15",
                                "Poggio et al. 2021",
                                "EcoDes-DK15",
                                "Poggio et al. 2021",
                                "EcoDes-DK15",
                                "Bjerreskov et al. 2021",
                                "EcoDes-DK15",
                                "EcoDes-DK15"),
       `Ecological Meaning` = c("Quality of lidar signal reflected (proxy of biomass).",
                                "Variation in quality of lidar signal reflected within 10 m pixel (proxy of variation in biomass).",
                                "Lidar estimator of canopy height (95-percentile of height distribution of all vegetation points in 10 m pixel).",
                                "Variation in lidar estimator of canopy height within 110 m focal window (11 x 11 pixels).",
                                "Estimated percentage clay content of soil (250 m resolution downscaled to 10 m).",
                                "Terrain height above sea level.",
                                "Variation in terrain height above sea level within 110 m focal window (11 x 11 pixels).",
                                "Foliage height diversity MacArthur and MacArthur (1979) based on height bins by Wilson (1974)",
                                "Estimated variation in canopy height within 10 m pixel.",
                                "Estimate of depth of near-surface groundwater during an average summer.",
                                "Variation in the estimate of depth of near-surface groundwater during an average summer within a 110 m focal window (11 x 11 pixels).",
                                "Presence of linear features in the terrain (valleys, ridges etc.) based on a 50 m search radius.",
                                "Estimated percentage sand content of soil (250 m resolution downscaled to 10 m).",
                                "Terrain slope at 10 m",
                                "Estimated percentage soil organic carbon content of soil (250 m resolution downscaled to 10 m).",
                                "Annual incident solar radiation based on terrain model (aspect and slope).",
                                "Decidous or coniferous forest.",
                                "Denisty of vegetation points in 10 m lidar pixel.",
                                "Variation of density of vegeation points amongst pixels within 110 m window (11 x 11 pixels)."
                                )) %>%
  kable() %>%
  kable_classic(full_width = F)  

```

## References
- Assmann, Jakob J., Jesper E. Moeslund, Urs A. Treier, and Signe Normand. "EcoDes-DK15: High-resolution ecological descriptors of vegetation and terrain derived from Denmark's national airborne laser scanning data set." Earth System Science Data Discussions (2021): 1-32.
-Bjerreskov, K. S., Nord-Larsen, T., and Fensholt, R.: Classification of Nemoral Forests with Fusion of Multi-Temporal Sentinel-1 and 2 Data, 13, 950, https://doi.org/10.3390/rs13050950, 2021.
- Hijmans, Robert J., Aniruddha Ghosh, and Alex Mandel. 2021. Geodata: Download Geographic Data. https://CRAN.R-project.org/package=geodata.
-Koch, J., Gotfredsen, J., Schneider, R., Troldborg, L., Stisen, S., and Henriksen, H. J.: High Resolution Water Table Modeling of the Shallow Groundwater Using a Knowledge-Guided Gradient Boosting Decision Tree Model, 3, 2021.
- MacArthur, R. H., & MacArthur, J. W. (1961). On Bird Species Diversity. Ecology, 42(3), 594–598. https://doi.org/10.2307/1932254
- Poggio, Laura, Luis M De Sousa, Niels H Batjes, Gerard Heuvelink, Bas Kempen, Eloi Ribeiro, and David
Rossiter. 2021. “SoilGrids 2.0: Producing Soil Information for the Globe with Quantified Spatial Uncertainty.” Soil 7 (1): 217–40.
- Willson, M. F. (1974). Avian Community Organization and Habitat Structure. Ecology, 55(5), 1017–1029.