---
title: "DK Forest LiDAR - Focal predictors selection"
author: "Jakob J. Assmann"
date: "2/3/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(kableExtra)
load("../data/training_data/pixel_training.Rda")
pixel_training_data <- pixel_training_data %>%
  st_drop_geometry() %>%
  select(contains("110"), contains("250"),
         dtm_10m,
         vegetation_density,
         ns_groundwater_summer,
         canopy_height)
pixel_training_data <- pixel_training_data[,order(colnames(pixel_training_data))]
```

## Content

We caluclated the mean and sd in 110 m and 250 m windows for the following variables:
1. dtm_10m
2. canopy_height
3. vegetation_density
4. ns_ground_water

Here is how those measures are correlated with their focal variables:

```{r correlations, echo = FALSE, results = "asis"}
# Generate correlation tables
res <- cor(pixel_training_data, use = "complete.obs") %>% 
  # Format numbers for looks
  (function(x) {
    x <- round(x, 2)
    x <- formatC(x, digits = 2, format = "f", flag = "+")
    x[lower.tri(x)] <- ""
    return(x)}) %>%
  # convert to dataframe
  as.data.frame() %>%
  # split by rows according to variable
  split(sort(rep(1:4, 5))) %>%
  # reduce columns to variable
  mapply(function(x,y) {
    x[,y]
    }, x = ., y = list(1:5,6:10,11:15,16:20),
    SIMPLIFY = F) %>%
  # map over each variable to create table
  lapply(function(x){
    cat("<strong>", rownames(x)[1], "</strong>\n")
    # tidy up row and column names
    rownames(x)[2:5] <- gsub(paste0(rownames(x)[1], "_?"), "", rownames(x)[2:5])
    colnames(x)[2:5] <- gsub(paste0(colnames(x)[1], "_?"), "", colnames(x)[2:5])
        rownames(x)[1] <- "cell_10m"
    colnames(x)[1] <- "cell_10m"
    # change background for cells with cor <= 0.5
    for(i in 1:ncol(x)){
      x[,i] <- cell_spec(x[,i], 
                         background = case_when(as.numeric(x[,i]) <= 0.5 ~ "lightgreen",
                                                TRUE ~ "white"))
    }
    # Plot out kable
    kable(x, format = "html", escape = F) %>%
      kable_classic(font_size = 12, full_width = F, position = "left") %>%
      print()
    # add some spaces to allow pandoc to recognise separation
    cat("\n\n\n\n")
    return(NULL)})
```

## Variation Inflation Factors

To reduce the number of features systematically, we calculate variance inflation
factors (vIFs). A VIF above 5 indicates that the variable introduces multicoliniearity
in the dataset. A conservative rule is to only keep variables with VIFs below 2.5.

Here we do a stepwise selection based on the VIFs and the correlation tables above. 
VIFs exceeding 5 are highlighted in red. 

### 1) All variables

``` {r echo = FALSE,}
vifs <- usdm::vif(pixel_training_data)
vifs$VIF <- cell_spec(round(vifs$VIF, 2),
                      background = case_when(vifs$VIF >= 5 ~ "red",
                                                TRUE ~ "white"),
                      color = case_when(vifs$VIF >= 5 ~ "white",
                                                TRUE ~ "black"))
kable(vifs, format = "html", escape = F) %>%
  kable_classic(font_size = 12, full_width = F, position = "left")

```

The mean variables seem to introduce a lot of colinearity (very high VIFs,
and see correlation tables above). We drop them first. 

### 2) Drop mean variables

``` {r echo = FALSE,}
vifs <- pixel_training_data %>% dplyr::select(-contains("mean")) %>% 
  usdm::vif() 
vifs$VIF <- cell_spec(round(vifs$VIF, 2),
                      background = case_when(vifs$VIF >= 5 ~ "red",
                                                TRUE ~ "white"),
                      color = case_when(vifs$VIF >= 5 ~ "white",
                                                TRUE ~ "black"))
kable(vifs, format = "html", escape = F) %>%
  kable_classic(font_size = 12, full_width = F, position = "left")
```

The focal variables of different window sizes are highly correlated with each
other. The correlation tables (above) suggest the 110 m windows are less correlated
with the 10 m cell values, so we drop the 250 m windows next. 

### 3) Drop 250 m variables

``` {r echo = FALSE,}
vifs <-  pixel_training_data %>% dplyr::select(-contains("mean"),
                                      -contains("250")) %>% 
  usdm::vif()
vifs$VIF <- cell_spec(round(vifs$VIF, 2),
                      background = case_when(vifs$VIF >= 5 ~ "red",
                                                TRUE ~ "white"),
                      color = case_when(vifs$VIF >= 5 ~ "white",
                                                TRUE ~ "black"))
kable(vifs, format = "html", escape = F) %>%
  kable_classic(font_size = 12, full_width = F, position = "left")

```

The final set of variables includes only the 10 m cell values and the sd calculated
for the 110 m windows. 