---
title: "DK Forest LiDAR - Gradien Boosting Model Performance"
author: "Jakob Assmann"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
library(sf)
library(caret)
library(tidyverse)
library(ggplot2)
library(ggtext)
library(rnaturalearth)
library(rasterVis)
library(kableExtra)
library(gbm)
```

## Training data overview

I generated a training dataset consisting of 200k pixel samples from the EcoDes-
DK15 grid. 100k samples each come from one of the two forest polygon sets 
("low" and "high" forest value). I then extracted the predictor data for the 
pixel centres for those data. 

``` {r training_data_map, echo = FALSE, cached = TRUE}
# Load pixel sample
load("../data/training_data/pixel_training.Rda")

# Get Denmark boundaries
denmark <- ne_countries(scale = "large", 
                        country = "denmark",
                        returnclass = "sf") %>%
  st_transform(st_crs(pixel_training_data))

# Plot training data
ggplot() +
  geom_sf(data = denmark, fill = "NA", colour = "black") +
  geom_sf(data = pixel_training_data, 
          mapping = aes(colour = forest_value),
          size = 0.01) +
  labs(title = "Pixel sample for training") +
  coord_sf(datum = st_crs(pixel_training_data)) +
  theme_void()
```

For the training I then split the data set into a training (80%) and validation 
partition (20%). This split was done based on two stratification. The BIOWIDE
stratification and Derek's stratification. The samples were split so that they 
were distributed evenly in each strata (80/20). 

I then trained GBM models on this (currently minimal hyperparameter tuning) using
the training dataset. This document evaluates the performance of each
stratification based on the validation dataset in both stratification.   

``` {r function_defs, echo = FALSE, cache=TRUE}

# Load regions
biowide_regions <- read_sf("../data/stratification/biowide_georegions/biowide_zones.shp")
dereks_strat <- rast("../data/stratification/derek_stratification/Results_2clim_5soil.tif") %>% as.polygons() %>% st_as_sf(crs = 25832) %>% mutate(Temp = factor(Temp))

# Function to Generate performance table from list of confusion matrices
performance_table <- function(cf_matrix_list){
  performance_stats <- lapply(cf_matrix_list, function(x){
    data.frame(Accuracy = round(x$overall["Accuracy"],2),
               Error = round(1 - x$overall["Accuracy"],2),
               Sensitivity = round(x$byClass["Sensitivity"],2),
               Specificity = round(x$byClass["Specificity"],2),
               FPR = round(1 - x$byClass["Specificity"],2),
               UserAccuracy = round(x$byClass["Pos Pred Value"],2))
  }) %>% bind_rows()
  performance_stats <- data.frame(t(performance_stats))
  names(performance_stats) <- names(cf_matrix_list)
  performance_stats$Measure <- c("Accuracy",
                                "Error",
                                "Sensitivity (True Positive Rate)",
                                "Specificity (True Negative Rate)",
                                "Fall-out (False Positive Rate)",
                                "Positive predictive value (User Accuracy)")
  row.names(performance_stats) <- NULL
  performance_stats <- relocate(performance_stats, Measure)
  return(performance_stats)
}

# Function to plot performance in biowide regions
plot_biowide_regions <- function(performance_overall,
                                            performance_list){
  
  # Extract statistics from performance_list
  row_order <- sapply(biowide_regions$region, function (x) which(x == names(performance_list)))
  biowide_regions$accuracy <- sapply(performance_list, 
                                     function(x) round(x$overall["Accuracy"],2))[row_order]
  biowide_regions$sens <- sapply(performance_list, 
                                 function(x) round(x$byClass["Sensitivity"],2))[row_order]
  biowide_regions$useracc <- sapply(performance_list, 
                                 function(x) round(x$byClass["Pos Pred Value"],2))[row_order]
  
  biowide_regions <- mutate(
    biowide_regions,
    x_cen = st_coordinates(st_centroid(biowide_regions))[,1],
    y_cen = st_coordinates(st_centroid(biowide_regions))[,2],
    map_span_x = st_bbox(biowide_regions)["xmax"]- 
      st_bbox(biowide_regions)["xmin"],
    map_span_y = st_bbox(biowide_regions)["ymax"]- 
      st_bbox(biowide_regions)["ymin"]) 
  
  biowide_regions <- 
    mutate(biowide_regions,
           x = x_cen + map_span_x * c(-0.10,  # Nordjlland
                                      -0.15,  # Vestjylland
                                      0.30,  # Oestjylland
                                      0.25,  # Sjaelland
                                      -0.00,  # Bornholm
                                      -0.20), # Fune_Lolland
           
           y = y_cen + map_span_y * c( 0.10,  # Nordjlland
                                       0.10,  # Vestjylland
                                       0.30,  # Oestjylland
                                       0.00,  # Sjaelland
                                       -0.10,  # Bornholm
                                       -0.20), # Fune_Lolland
           hjust = c(1, # Nordjlland
                     1, # Vestjylland
                     0, # Oestjylland
                     0, # Sjaelland
                     0, # Bornholm
                     0),# Fune_Lolland
           vjust = c(0, # Nordjlland
                     1, # Vestjylland
                     0, # Oestjylland
                     0, # Sjaelland
                     1, # Bornholm
                     1) # Fune_Lolland
    )
  
  biowide_regions <- mutate(
    biowide_regions,
    label = paste0("<span style='font-size:16pt'>",
                   "**", region, "**",
                   "</span>",
                   "<span style='font-size:10pt; color:black'>",
                   "<br>Accuracy: ", accuracy,
                   "<br>Sensitivity: ", sens,
                   "<br>User Accuracy: ", useracc,
                   "</span>"))
  ggplot() + 
    geom_sf(aes(fill = region), data = biowide_regions) +
    geom_richtext(aes(x = x, 
                      y = y, 
                      label = label,
                      colour = region,
                      hjust = hjust,
                      vjust = vjust), 
                  data = biowide_regions,
                  fill = NA,
                  label.color = NA) +
    coord_sf(clip = "off",
             crs = st_crs(biowide_regions),
             xlim = st_bbox(biowide_regions)[c(1,3)] * c(0.7, 1.15),
             ylim = st_bbox(biowide_regions)[c(2,4)] * c(0.975, 1.0125)) +
    labs(title = "Overall Performance",
      subtitle = paste0(
      "Accuracy: ", round(performance_overall$overall["Accuracy"], 2), 
      "\nSensitivity: ", round(performance_overall$byClass["Sensitivity"], 2),
      "\nUser Accuracy: ", round(performance_overall$byClass["Pos Pred Value"], 2))) +
    theme_void() +
    theme(legend.position = "none",,
          plot.margin = margin(0,0,0,0, unit = "in"))
}

# Function to plot performance in Derek's regions
plot_derek_regions <- function(performance_overall,
                                 performance_list){
  dereks_regions <- data.frame(region = paste0("Region ", c(1,2,3)))
  names(performance_list) <- paste0("Region ", names(performance_list))
  # Extract statistics from performance_list
  row_order <- sapply(dereks_regions$region, function (x) which(x == names(performance_list)))
  dereks_regions$accuracy <- sapply(performance_list, 
                                     function(x) round(x$overall["Accuracy"],2))[row_order]
  dereks_regions$sens <- sapply(performance_list, 
                                 function(x) round(x$byClass["Sensitivity"],2))[row_order]
  dereks_regions$useracc <- sapply(performance_list, 
                                 function(x) round(x$byClass["Pos Pred Value"],2))[row_order]
  
  dereks_regions <- 
    mutate(dereks_regions,
           x = c(567108.7, 567108.7, 567108.7) + sum(-309034, 1027089) * 
             c(0.3, -0.1, -0.2), 
           y = c(6201086, 6201086, 6201086) + sum(-5974061, 6482333) *
             c(0.25, 0.3, 0.1),
           hjust = c(0,1,1),
           vjust = c(1,0,1)
    )
  
  dereks_regions <- mutate(
    dereks_regions,
    label = paste0("<span style='font-size:16pt'>",
                   "**", region, "**",
                   "</span>",
                   "<span style='font-size:10pt; color:black'>",
                   "<br>Accuracy: ", accuracy,
                   "<br>Sensitivity: ", sens,
                   "<br>User Accuracy: ", useracc,
                   "</span>"))
  ggplot() + 
    geom_sf(data = dereks_strat, mapping = aes(fill = Temp), colour = NA) +
    geom_richtext(aes(x = x, 
                      y = y, 
                      label = label,
                      colour = region,
                      hjust = hjust,
                      vjust = vjust), 
                  data = dereks_regions,
                  fill = NA,
                  label.color = NA) +
    coord_sf(clip = "off",
             crs = st_crs(25832),
             xlim = c(309034, 1027089),
             ylim = c(5898440 , 6482333)) +
    labs(title = "Overall Performance",
      subtitle = paste0(
        "Accuracy: ", round(performance_overall$overall["Accuracy"], 2), 
        "\nSensitivity: ", round(performance_overall$byClass["Sensitivity"], 2),
        "\nUser Accuracy: ", round(performance_overall$byClass["Pos Pred Value"], 2))) +
    theme_void() +
    theme(legend.position = "none",,
          plot.margin = margin(0,0,0,0, unit = "in"))
}

```

## Models trained using BIOWIDE stratification

For these models the training data was split according to the BIOWIDE 
stratification.

```{r data_prep_biowide, echo = FALSE, results='hide'}
# Load validation data
load("../data/training_data/pixel_valid_biowide.Rda")
# Load training data
load("../data/training_data/pixel_training_biowide.Rda")

# Load model
load("../data/models/final_gbm_model_pixel_biowide.Rda")
gbm_biowide <- gbm_fit
rm(gbm_fit)

# Calculate performance stats for validation data
performance_biowide_overall <- predict(gbm_biowide, 
                                       newdata = pixel_valid_biowide) %>%
  confusionMatrix(data = ., 
                  pixel_valid_biowide$forest_value)

performance_biowide_biowide <- pixel_valid_biowide %>% 
  ungroup() %>%
  split(., .$biowide_region) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_biowide, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

performance_biowide_derek <- pixel_valid_biowide %>% 
  ungroup() %>%
  split(., .$dereks_stratification) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_biowide, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

performance_biowide_forest_type <- pixel_valid_biowide %>% 
  ungroup() %>%
  split(., mutate(., tree_type = case_when(treetype_bjer_dec == 0 ~ "Coniferous",
                                           treetype_bjer_dec == 1 ~ "Broadleaf")) %>%
          pull(tree_type)) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_biowide, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

# Calculate performance stats for training data
performance_biowide_overall_training <- predict(gbm_biowide, 
                                       newdata = pixel_training_biowide) %>%
  confusionMatrix(data = ., 
                  pixel_training_biowide$forest_value)

performance_biowide_biowide_training <- pixel_training_biowide %>% 
  ungroup() %>%
  split(., .$biowide_region) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_biowide, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

performance_biowide_derek_training <- pixel_training_biowide %>% 
  ungroup() %>%
  split(., .$dereks_stratification) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_biowide, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

performance_biowide_forest_type_training <- pixel_training_biowide %>% 
  ungroup() %>%
  split(., mutate(., tree_type = case_when(treetype_bjer_dec == 0 ~ "Coniferous",
                                           treetype_bjer_dec == 1 ~ "Broadleaf")) %>%
          pull(tree_type)) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_biowide, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

```

### Variable importance

Variable importance for this boosted regression tree model.

``` {r biowide_variable_importnace, echo = FALSE}
summary(gbm_biowide, plotit = F)
```

### Performance in BIOWIDE regions:

Performance map based on the <span style="color: green">independent validation data</span>:

``` {r biowide_biowide_performance_map, echo = FALSE}
suppressWarnings(
  plot_biowide_regions(performance_biowide_overall,
                       performance_biowide_biowide))
```

Performance table based on the <span style="color: green">independent validation data</span>:

``` {r biowide_biowide_performance_table, echo = FALSE}
performance_table(setNames(c(list(performance_biowide_overall), performance_biowide_biowide), 
                           c("Overall", names(performance_biowide_biowide)))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

Performance table based on the <span style="color: red">dependent training data</span>:

``` {r biowide_biowide_performance_table_training, echo = FALSE}
performance_table(setNames(c(list(performance_biowide_overall_training), performance_biowide_biowide_training), 
                           c("Overall", names(performance_biowide_biowide_training)))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

### Perfromance in Derek's regions:

Performance map based on the <span style="color: green">independent validation data</span>:

``` {r biowide_derek_performance_map, echo = FALSE}
suppressWarnings(
  plot_derek_regions(performance_biowide_overall,
                     performance_biowide_derek))
```

Performance table based on the <span style="color: green">independent validation data</span>:

``` {r biowide_derek_performance_table, echo = FALSE}
performance_table(setNames(c(list(performance_biowide_overall), performance_biowide_derek), 
                  c("Overall", paste0("Region ", names(performance_biowide_derek))))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

Performance table based on the <span style="color: red">dependent training data</span>:

``` {r biowide_derek_performance_table_training, echo = FALSE}
performance_table(setNames(c(list(performance_biowide_overall_training), performance_biowide_derek_training), 
                  c("Overall", paste0("Region ", names(performance_biowide_derek_training))))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

### Performance by forest type (boradleaf vs. coniferous)

Performance table based on the <span style="color: green">independent validation data</span>:

``` {r biowide_performance_forest_type, echo = FALSE}
performance_table(setNames(c(list(performance_biowide_overall), performance_biowide_forest_type), 
                  c("Overall", names(performance_biowide_forest_type)))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

Performance table based on the <span style="color: red">dependent training data</span>:

``` {r biowide_performance_forest_type_training, echo = FALSE}
performance_table(setNames(c(list(performance_biowide_overall_training), performance_biowide_forest_type_training), 
                  c("Overall", names(performance_biowide_forest_type_training)))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

## Models trained using Derek's stratification

``` {r data_prep_derek, echo = FALSE, results='hide'}
# Load validation data
load("../data/training_data/pixel_valid_derek.Rda")

# Load training data
load("../data/training_data/pixel_training_derek.Rda")

# Load model
load("../data/models/final_gbm_model_pixel_derek.Rda")
gbm_derek <- gbm_fit
rm(gbm_fit)

# Calculate performance statistics for validation data
performance_derek_overall <- predict(gbm_derek, 
                                       newdata = pixel_valid_derek) %>%
  confusionMatrix(data = ., 
                  pixel_valid_derek$forest_value)

performance_derek_biowide <- pixel_valid_derek %>% 
  ungroup() %>%
  split(., .$biowide_region) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_derek, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

performance_derek_derek <- pixel_valid_derek %>% 
  ungroup() %>%
  split(., .$dereks_stratification) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_derek, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  })

performance_derek_forest_type <- pixel_valid_derek %>% 
  ungroup() %>%
  split(., mutate(., tree_type = case_when(treetype_bjer_dec == 0 ~ "Coniferous",
                                           treetype_bjer_dec == 1 ~ "Broadleaf")) %>%
          pull(tree_type)) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_biowide, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

# Calculate performance statistics for training data
performance_derek_overall_training <- predict(gbm_derek, 
                                       newdata = pixel_training_derek) %>%
  confusionMatrix(data = ., 
                  pixel_training_derek$forest_value)

performance_derek_biowide_training <- pixel_training_derek %>% 
  ungroup() %>%
  split(., .$biowide_region) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_derek, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 

performance_derek_derek_training <- pixel_training_derek %>% 
  ungroup() %>%
  split(., .$dereks_stratification) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_derek, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  })

performance_derek_forest_type_training <- pixel_training_derek %>% 
  ungroup() %>%
  split(., mutate(., tree_type = case_when(treetype_bjer_dec == 0 ~ "Coniferous",
                                           treetype_bjer_dec == 1 ~ "Broadleaf")) %>%
          pull(tree_type)) %>%
  lapply(function(test_data_region){
    test_preds_region <- predict(gbm_biowide, newdata = test_data_region)
    #cat(unique(test_data_region$biowide_region), "\n")
    confusionMat <- confusionMatrix(data = test_preds_region, test_data_region$forest_value)
    #print(confusionMat)
    return(confusionMat)
  }) 
```

### Variable importance

Variable importance for this boosted regression tree model.

```{r derek_variable_importance, echo = FALSE}
summary(gbm_derek, plotit = F)
```

### Performance in BIOWIDE regions:

Performance map based on the <span style="color: green">independent validation data</span>:

```{r derek_biowide_performance_map, echo = FALSE}
suppressWarnings(
  plot_biowide_regions(performance_derek_overall,
                       performance_derek_biowide))
```

Performance table based on the <span style="color: green">independent validation data</span>:

``` {r derek_biowide_performance_table, echo = FALSE}
performance_table(setNames(c(list(performance_derek_overall), performance_derek_biowide), 
                  c("Overall", names(performance_derek_biowide)))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

Performance table based on the <span style="color: red">dependent training data</span>:

``` {r derek_biowide_performance_table_training, echo = FALSE}
performance_table(setNames(c(list(performance_derek_overall_training), performance_derek_biowide_training), 
                  c("Overall", names(performance_derek_biowide_training)))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

### Perfromance in Derek's regions:

Performance map based on the <span style="color: green">independent validation data</span>:

``` {r derek_derek_performance_map, echo = FALSE}
suppressWarnings(
  plot_derek_regions(performance_derek_overall,
                     performance_derek_derek))
```

Performance table based on the <span style="color: green">independent validation data</span>:

``` {r derek_derek_performance_table, echo = FALSE}
performance_table(setNames(c(list(performance_derek_overall), performance_derek_derek), 
                  c("Overall", paste0("Region ", names(performance_derek_derek))))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

Performance table based on the <span style="color: red">dependent training data</span>:

``` {r derek_derek_performance_table_training, echo = FALSE}
performance_table(setNames(c(list(performance_derek_overall_training), performance_derek_derek_training), 
                  c("Overall", paste0("Region ", names(performance_derek_derek_training))))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

### Performance by forest type (boradleaf vs. coniferous)

Performance table based on the <span style="color: green">independent validation data</span>:

``` {r derek_performance_forest_type, echo = FALSE}
performance_table(setNames(c(list(performance_derek_overall), performance_derek_forest_type), 
                  c("Overall", names(performance_biowide_forest_type)))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```

Performance table based on the <span style="color: red">dependent training data</span>:

``` {r derek_performance_forest_type_training, echo = FALSE}
performance_table(setNames(c(list(performance_derek_overall_training), performance_derek_forest_type_training), 
                  c("Overall", names(performance_biowide_forest_type_training)))) %>%
  kable() %>% kable_classic(font_size = 12, full_width = F, position = "left")
```
