---
title: "DK Forest LiDAR Summary Stats for Projections"
author: "Jakob Johann Assmann"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
library(sf)
library(tidyverse)
library(ggplot2)
library(kableExtra)
```

This document provides summary stats for the forest quality projections. Here, 
we concentrate on the models (gradient boosting and random forests) trained with 
the "BIOWIDE" stratification as these performed best overall. 

Content:

1. Forest area in Denmark according to Basemap 03.
2. Disturbance detected in forests overall.
2. Gradient Boosting model summary tats.
3. Random Forest model summary stats.

## Forests in Denmark according to Basemap 03

The forest mask used for our projections is based on the DCE Basemap 03 
sub-layer "tree cover" for 2016 [(Levin 2019)](https://dce2.au.dk/pub/TR159.pdf).

The sub-layer contains five "object types": 1) tree cover, 2) forest / afforestation, 
3) Christmas trees / cut greenery, 4) nursery / plantation, and 5) energy forest. 

The table below shows how much area each of the classes cover in the layer (see 
also Table 4.3, Levin 2019):

``` {r basemap_forest_area, echo = FALSE}
tibble(
  Code = c(as.character(1:5), "-"),
  Name = c("tree cover", 
           "forest / afforestation", 
           "Christmas trees / cut greenery",
           "nursery / plantation", 
           "energy forest",
           "total"),
  `Area [km²]` = c(928.3,
                   5633.9,
                   176.2,
                   46.8,
                   91.4,
                   6876.5),
  `Proportion [%]` = c(
    13.5,
    81.9,
    2.6,
    0.7,
    1.3,
    100
  )
) %>% 
  kable() %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

For our projections **we only use the "forest / afforestation" layer (2)**. 

To match the grid of the EcoDes-DK15 rasters we had to project the forest mask.
For this we used a nearest neighbour algorithm. Here we simply confirm that the 
forest area (code 2) in the final mask "forest_mask.tif" matches the area noted 
in the table above. 

```{r area_forest_mask, echo = FALSE, cache = TRUE}
forest_mask <- rast("../data/basemap_forests/forest_mask.tif")
total_forest_area <- (global(forest_mask == 1, "sum", na.rm = T) * (10 * 10))$sum[1]
tibble(
  `Layer` = "forest mask",
  `Area [km²]` = round(total_forest_area / 1000^2, 2)
) %>% 
  kable() %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Disturbance overall 

We used a disturbance layer generated by Cornelius [(Senf and Seidl 2021)](https://zenodo.org/record/4746129) 
to estimate the disturbance in Denmark's forests since the lidar data for EcoDes-DK15 was collected. 

Please note that this disturbance mask was projected and down-sampled from a 30 m Landsat grid to the 10 m EcoDes-DK15 grid (nearest 
neighbour algorithm), potentially adding small uncertainties to the area estimates. Currently, we also only account for disturbances from 2016 till 2020. 


``` {r disturbance_stats, echo = FALSE, cache = TRUE}
disturbance_mask <- rast("../data/forest_change_cs/disturbance_since_2015.tif")
gbm_biowide_projections <- rast("../data/projections/gbm_biowide/forest_quality_gbm_biowide.vrt")
disturbance_mask <- mask(disturbance_mask, gbm_biowide_projections)
disturbance_mask_area <- (global(disturbance_mask, "sum", na.rm = T) * (10 * 10))$sum[1]
tibble(
  Name = c("disturbed forest", "total forest"),
  `Area [km²]` = round(c(disturbance_mask_area / 1000^2, 
                         total_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * disturbance_mask_area / total_forest_area, 
                           100), 1), digits = 2, format = "f")
) %>% 
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Gradient Boosting projections summary stats

This gradient boosting model was trained based on the "BIOWIDE" stratification.

``` {r gbm_biowide_stats, echo = FALSE, cache = TRUE}
high_quality_forest <- gbm_biowide_projections == 1
high_quality_forest_area <- (global(high_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest <- gbm_biowide_projections == 2
low_quality_forest_area <- (global(low_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("high quality forest",
             "low quality forest",
             "total forest"),
  `Area [km²]` = round(c(high_quality_forest_area / 1000^2,
                         low_quality_forest_area / 1000^2,
                         total_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_area / total_forest_area,
                                     100 * low_quality_forest_area / total_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

Disturbance statistics:

``` {r gbm_biowide_disturbance, echo = FALSE, cache = TRUE}
disturbance_mask_applied <- disturbance_mask
disturbance_mask_applied[disturbance_mask_applied != 1] <- NA
high_quality_forest_disturbed <- mask(high_quality_forest, disturbance_mask_applied)
high_quality_forest_disturbed_area <- (global(high_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest_disturbed <- mask(low_quality_forest, disturbance_mask_applied) 
low_quality_forest_disturbed_area <- (global(low_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("disturbed high quality forest",
             "total high quality forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         high_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / high_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed low quality forest",
             "total low quality forest"),
  `Area [km²]` = round(c(low_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * low_quality_forest_disturbed_area / low_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed high quality forest",
             "disturbed low quality forest",
             "total disturbed forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_disturbed_area / 1000^2,
                         disturbance_mask_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / disturbance_mask_area,
                                     100 * low_quality_forest_disturbed_area / disturbance_mask_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Random Forest projections summary stats

This random forest model was trained based on the "BIOWIDE" stratification.

``` {r ranger_biowide_stats, echo = FALSE, cache = TRUE}
ranger_biowide_projections <- rast("../data/projections/ranger_biowide/forest_quality_ranger_biowide.vrt")
high_quality_forest <- ranger_biowide_projections == 1
high_quality_forest_area <- (global(high_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest <- ranger_biowide_projections == 2
low_quality_forest_area <- (global(low_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("high quality forest",
             "low quality forest",
             "total forest"),
  `Area [km²]` = round(c(high_quality_forest_area / 1000^2,
                         low_quality_forest_area / 1000^2,
                         total_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_area / total_forest_area,
                                     100 * low_quality_forest_area / total_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

Disturbance statistics:

``` {r ranger_biowide_disturbance, echo = FALSE, cache = TRUE}
disturbance_mask_applied <- disturbance_mask
disturbance_mask_applied[disturbance_mask_applied != 1] <- NA
high_quality_forest_disturbed <- mask(high_quality_forest, disturbance_mask_applied)
high_quality_forest_disturbed_area <- (global(high_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest_disturbed <- mask(low_quality_forest, disturbance_mask_applied) 
low_quality_forest_disturbed_area <- (global(low_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("disturbed high quality forest",
             "total high quality forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         high_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / high_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed low quality forest",
             "total low quality forest"),
  `Area [km²]` = round(c(low_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * low_quality_forest_disturbed_area / low_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed high quality forest",
             "disturbed low quality forest",
             "total disturbed forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_disturbed_area / 1000^2,
                         disturbance_mask_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / disturbance_mask_area,
                                     100 * low_quality_forest_disturbed_area / disturbance_mask_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```
