---
title: "DK Forest LiDAR Summary Stats for Projections"
author: "Jakob Johann Assmann"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
library(sf)
library(tidyverse)
library(ggplot2)
library(kableExtra)
```

This document provides summary stats (area) for the forest conservation value projections.
We show the statistics for all four models tested in our analysis. 

Content:

<!-- 1. Forest area in Denmark according to Basemap 03. -->
1. Forest area in Denmark according to Bjerreskov et al. 2021.
2. Training data area: High conservation value and low conservation value forests.
3. Disturbance detected in forests overall.
4. Gradient Boosting model summary stats (BIOWIDE).
5. Random Forest model summary stats (BIOWIDE).
6. Gradient Boosting model summary stats (Derek's stratification).
7. Random Forest model summarz stats (Derek's stratification).

<!-- ## Forests in Denmark according to Basemap 03 -->

<!-- The forest mask used for our projections is based on the DCE Basemap 03  -->
<!-- sub-layer "tree cover" for 2016 [(Levin 2019)](https://dce2.au.dk/pub/TR159.pdf). -->

<!-- The sub-layer contains five "object types": 1) tree cover, 2) forest / afforestation,  -->
<!-- 3) Christmas trees / cut greenery, 4) nursery / plantation, and 5) energy forest.  -->

<!-- The table below shows how much area each of the classes cover in the layer (see  -->
<!-- also Table 4.3, Levin 2019): -->

<!-- ``` {r basemap_forest_area, echo = FALSE} -->
<!-- tibble( -->
<!--   Code = c(as.character(1:5), "-"), -->
<!--   Name = c("tree cover",  -->
<!--            "forest / afforestation",  -->
<!--            "Christmas trees / cut greenery", -->
<!--            "nursery / plantation",  -->
<!--            "energy forest", -->
<!--            "total"), -->
<!--   `Area [km²]` = c(928.3, -->
<!--                    5633.9, -->
<!--                    176.2, -->
<!--                    46.8, -->
<!--                    91.4, -->
<!--                    6876.5), -->
<!--   `Proportion [%]` = c( -->
<!--     13.5, -->
<!--     81.9, -->
<!--     2.6, -->
<!--     0.7, -->
<!--     1.3, -->
<!--     100 -->
<!--   ) -->
<!-- ) %>%  -->
<!--   kable() %>% -->
<!--   kable_classic(position = "left", -->
<!--                 full_width = FALSE) -->
<!-- ``` -->

<!-- For our projections **we only use the "forest / afforestation" layer (2)**.  -->

<!-- To match the grid of the EcoDes-DK15 rasters we had to project the forest mask. -->
<!-- For this we used a nearest neighbour algorithm. Here we simply confirm that the  -->
<!-- forest area (code 2) in the final mask "forest_mask.tif" matches the area noted  -->
<!-- in the table above.  -->

<!-- ```{r area_forest_mask, echo = FALSE, cache = FALSE} -->
<!-- forest_mask <- rast("../data/basemap_forests/forest_mask.tif") -->
<!-- total_forest_area <- (global(forest_mask == 1, "sum", na.rm = T) * (10 * 10))$sum[1] -->
<!-- tibble( -->
<!--   `Layer` = "forest mask", -->
<!--   `Area [km²]` = round(total_forest_area / 1000^2, 2) -->
<!-- ) %>%  -->
<!--   kable() %>% -->
<!--   kable_classic(position = "left", -->
<!--                 full_width = FALSE) -->
<!-- ``` -->

## Forests area in Denmark according to Bjerreskov et al. 2021

Below you can find the total area of forest in the forest mask from Bjerreskov et al. 2021.
This is the reference for the total area of forest used in our project. The mask is
based on the tree type layer from the same publication (see predictor description). We generated
the forest mask by refining the treetype layer into a forest mask by applying a minimum mapping filter removing all continuous forest patches smaller than 500 m2 (see also Bjerreskov et al. 2021).

```{r area_forest_mask, echo = FALSE, cache = FALSE}
forest_mask <- rast("../data/predictor_data/treetype/forest_mask_bjer_above_half_ha.tif")
total_forest_area <- (global(forest_mask == 1, "sum", na.rm = T) * (10 * 10))$sum[1]
tibble(
  `Layer` = "forest mask",
  `Area [km²]` = round(total_forest_area / 1000^2, 2)
) %>%
  kable() %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Training data area: High conservation value and low conservation value forests

Here you can see the area covered by our training data. Including both the
high conservation value forests with designations (p15, p25 and private old growth), as well as the 
low conservation value training polygons. The proportions are given relative to the total area of
forest according to the forest mask generated from Bjerreskov et al. 2021 (see above). 

```{r designations_area, echo = FALSE}
load("../data/training_data/polygon_geometries/high_quality_polys.Rda")
load("../data/training_data/polygon_geometries/low_quality_polys.Rda")

tibble(
  category = c(unique(high_quality$polygon_source),
               "total_high_quality",
               unique(low_quality$polygon_source),
               "total_low_quality"),
  `Area [km²]` = round(c(high_quality %>% 
                           group_by(polygon_source) %>%
                           summarize() %>%
                           st_area(),
                         high_quality%>%
                           summarize() %>%
                           st_area(),
                         low_quality %>% 
                           group_by(polygon_source) %>%
                           summarize() %>%
                           st_area(),
                         low_quality %>%
                           summarize() %>%
                           st_area()) / 1000^2, 2) %>% 
    as.numeric()
) %>% 
  mutate(`Proportion of all forest [%]` = round((`Area [km²]` / total_forest_area) * 100, 6)) %>%
  kable() %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Disturbance overall 

We used a disturbance layer generated by Cornelius [(Senf and Seidl 2021)](https://zenodo.org/record/4746129) 
to estimate the disturbance in Denmark's forests since the lidar data for EcoDes-DK15 was collected. 

Please note that this disturbance mask was projected and down-sampled from a 30 m Landsat grid to the 10 m EcoDes-DK15 grid (nearest 
neighbour algorithm), potentially adding small uncertainties to the area estimates. Currently, we also only account for disturbances from 2016 till 2020. 


``` {r disturbance_stats, echo = FALSE, cache = FALSE}
disturbance_mask <- rast("../data/forest_change_cs/disturbance_since_2015.tif")
gbm_biowide_projections <- rast("../data/projections/gbm_biowide/forest_quality_gbm_biowide.vrt")
disturbance_mask <- mask(disturbance_mask, gbm_biowide_projections)
disturbance_mask_area <- (global(disturbance_mask, "sum", na.rm = T) * (10 * 10))$sum[1]
tibble(
  Name = c("disturbed forest", "total forest"),
  `Area [km²]` = round(c(disturbance_mask_area / 1000^2, 
                         total_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * disturbance_mask_area / total_forest_area, 
                           100), 1), digits = 2, format = "f")
) %>% 
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Gradient Boosting projections summary stats (BIOWIDE)

This gradient boosting model was trained based on the "BIOWIDE" stratification.

``` {r gbm_biowide_stats, echo = FALSE, cache = FALSE}
high_quality_forest <- gbm_biowide_projections == 1
high_quality_forest_area <- (global(high_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest <- gbm_biowide_projections == 2
low_quality_forest_area <- (global(low_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("high conservation value forest",
             "low conservation value forest",
             "total forest"),
  `Area [km²]` = round(c(high_quality_forest_area / 1000^2,
                         low_quality_forest_area / 1000^2,
                         total_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_area / total_forest_area,
                                     100 * low_quality_forest_area / total_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

Disturbance statistics:

``` {r gbm_biowide_disturbance, echo = FALSE, cache = FALSE}
disturbance_mask_applied <- disturbance_mask
disturbance_mask_applied[disturbance_mask_applied != 1] <- NA
high_quality_forest_disturbed <- mask(high_quality_forest, disturbance_mask_applied)
high_quality_forest_disturbed_area <- (global(high_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest_disturbed <- mask(low_quality_forest, disturbance_mask_applied) 
low_quality_forest_disturbed_area <- (global(low_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("disturbed high conservation value forest",
             "total high conservation value forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         high_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / high_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed low conservation value forest",
             "total low conservation value forest"),
  `Area [km²]` = round(c(low_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * low_quality_forest_disturbed_area / low_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed high conservation value forest",
             "disturbed low conservation value forest",
             "total disturbed forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_disturbed_area / 1000^2,
                         disturbance_mask_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / disturbance_mask_area,
                                     100 * low_quality_forest_disturbed_area / disturbance_mask_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Random Forest projections summary stats (BIOWIDE)

This random forest model was trained based on the "BIOWIDE" stratification.

``` {r ranger_biowide_stats, echo = FALSE, cache = FALSE}
ranger_biowide_projections <- rast("../data/projections/ranger_biowide/forest_quality_ranger_biowide.vrt")
high_quality_forest <- ranger_biowide_projections == 1
high_quality_forest_area <- (global(high_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest <- ranger_biowide_projections == 2
low_quality_forest_area <- (global(low_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("high conservation value forest",
             "low conservation value forest",
             "total forest"),
  `Area [km²]` = round(c(high_quality_forest_area / 1000^2,
                         low_quality_forest_area / 1000^2,
                         total_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_area / total_forest_area,
                                     100 * low_quality_forest_area / total_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

Disturbance statistics:

``` {r ranger_biowide_disturbance, echo = FALSE, cache = FALSE}
disturbance_mask_applied <- disturbance_mask
disturbance_mask_applied[disturbance_mask_applied != 1] <- NA
high_quality_forest_disturbed <- mask(high_quality_forest, disturbance_mask_applied)
high_quality_forest_disturbed_area <- (global(high_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest_disturbed <- mask(low_quality_forest, disturbance_mask_applied) 
low_quality_forest_disturbed_area <- (global(low_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("disturbed high conservation value forest",
             "total high conservation value forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         high_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / high_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed low conservation value forest",
             "total low conservation value forest"),
  `Area [km²]` = round(c(low_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * low_quality_forest_disturbed_area / low_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed high conservation value forest",
             "disturbed low conservation value forest",
             "total disturbed forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_disturbed_area / 1000^2,
                         disturbance_mask_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / disturbance_mask_area,
                                     100 * low_quality_forest_disturbed_area / disturbance_mask_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Gradient Boosting projections summary stats (Derek's stratification)

This gradient boosting model was trained based on the "Derek's" stratification.

``` {r gbm_derek_stats, echo = FALSE, cache = FALSE}
gbm_derek_projections <- rast("../data/projections/gbm_derek/forest_quality_gbm_derek.vrt")
high_quality_forest <- gbm_derek_projections == 1
high_quality_forest_area <- (global(high_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest <- gbm_derek_projections == 2
low_quality_forest_area <- (global(low_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("high conservation value forest",
             "low conservation value forest",
             "total forest"),
  `Area [km²]` = round(c(high_quality_forest_area / 1000^2,
                         low_quality_forest_area / 1000^2,
                         total_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_area / total_forest_area,
                                     100 * low_quality_forest_area / total_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

Disturbance statistics:

``` {r gbm_derek_disturbance, echo = FALSE, cache = FALSE}
disturbance_mask <- rast("../data/forest_change_cs/disturbance_since_2015.tif")
gbm_biowide_projections <- rast("../data/projections/gbm_biowide/forest_quality_gbm_biowide.vrt")
disturbance_mask <- mask(disturbance_mask, gbm_biowide_projections)
disturbance_mask_applied <- disturbance_mask
disturbance_mask_applied[disturbance_mask_applied != 1] <- NA
high_quality_forest_disturbed <- mask(high_quality_forest, disturbance_mask_applied)
high_quality_forest_disturbed_area <- (global(high_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest_disturbed <- mask(low_quality_forest, disturbance_mask_applied) 
low_quality_forest_disturbed_area <- (global(low_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("disturbed high conservation value forest",
             "total high conservation value forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         high_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / high_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed low conservation value forest",
             "total low conservation value forest"),
  `Area [km²]` = round(c(low_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * low_quality_forest_disturbed_area / low_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed high conservation value forest",
             "disturbed low conservation value forest",
             "total disturbed forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_disturbed_area / 1000^2,
                         disturbance_mask_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / disturbance_mask_area,
                                     100 * low_quality_forest_disturbed_area / disturbance_mask_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

## Random Forest projections summary stats (Derek's stratification)

This random forest model was trained based on the "Derek's" stratification.

``` {r ranger_derek_stats, echo = FALSE, cache = FALSE}
ranger_derek_projections <- rast("../data/projections/ranger_derek/forest_quality_ranger_derek.vrt")
high_quality_forest <- ranger_derek_projections == 1
high_quality_forest_area <- (global(high_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest <- ranger_derek_projections == 2
low_quality_forest_area <- (global(low_quality_forest, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("high conservation value forest",
             "low conservation value forest",
             "total forest"),
  `Area [km²]` = round(c(high_quality_forest_area / 1000^2,
                         low_quality_forest_area / 1000^2,
                         total_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_area / total_forest_area,
                                     100 * low_quality_forest_area / total_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```

Disturbance statistics:

``` {r ranger_derek_disturbance, echo = FALSE, cache = FALSE}
disturbance_mask <- rast("../data/forest_change_cs/disturbance_since_2015.tif")
gbm_biowide_projections <- rast("../data/projections/gbm_biowide/forest_quality_gbm_biowide.vrt")
disturbance_mask <- mask(disturbance_mask, gbm_biowide_projections)
disturbance_mask_applied <- disturbance_mask
disturbance_mask_applied[disturbance_mask_applied != 1] <- NA
high_quality_forest_disturbed <- mask(high_quality_forest, disturbance_mask_applied)
high_quality_forest_disturbed_area <- (global(high_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]
low_quality_forest_disturbed <- mask(low_quality_forest, disturbance_mask_applied) 
low_quality_forest_disturbed_area <- (global(low_quality_forest_disturbed, "sum", na.rm = T) * (10 * 10))$sum[1]

tibble(
  `Type` = c("disturbed high conservation value forest",
             "total high conservation value forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         high_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / high_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed low conservation value forest",
             "total low conservation value forest"),
  `Area [km²]` = round(c(low_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * low_quality_forest_disturbed_area / low_quality_forest_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)

tibble(
  `Type` = c("disturbed high conservation value forest",
             "disturbed low conservation value forest",
             "total disturbed forest"),
  `Area [km²]` = round(c(high_quality_forest_disturbed_area / 1000^2,
                         low_quality_forest_disturbed_area / 1000^2,
                         disturbance_mask_area / 1000^2), 2),
  `Proportion [%]` = formatC(round(c(100 * high_quality_forest_disturbed_area / disturbance_mask_area,
                                     100 * low_quality_forest_disturbed_area / disturbance_mask_area,
                                     100), 1), digits = 2, format = "f")
) %>%
  kable(align = c('l','r','r')) %>%
  kable_classic(position = "left",
                full_width = FALSE)
```
