---
title: "DK Forest LiDAR - Forest Annotations & Training Data"
author: Jakob J. Assmann
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(sf)
library(tidyverse)
library(ggplot2)
library(terra)
library(rnaturalearth)
```

This document provides an overview of the forest annotations used for generating the 
training dataset that forms the base of our forest quality models for Denmark. 

These annotations are vector polygons of forests in Denmark that are of known 
"high" or "low" quality. We used these polygons to generate a training dataset 
of 60k pixels based on the 10 m grid of Denmark that is used by our models. 
The grid is defined by the EcoDes-DK15 dataset. A brief description on how the
final pixel training dataset was generated from the forest 
annotations can be found at the end of this document. 

Note: What makes a "high" or "low" quality forest is to some degree arbitrary. 
Our definitions here are the result of a long discussion and have been developed over
multiple years. The aim was to arrive at a workable definition that aligns 
with the current framework of forest designations in Denmark, while also ensuring that 
enough training data is available. We appreciate that our chosen definitions of 
forest quality are a simplification and not without flaws (e.g., we assume that 
all "plantations" are of low forest quality). 

---

## High quality forests

Total number of high quality forest polygons: 9400.

### §25 and §15 forest

The core of the high quality forest annotations is made up by the polygons for 
the designated §25 ("naturmaessigt saerlig vaerdifuld skov") and §15 ("skovnatur") forests. 
The vector boundaries of these forests were retrieved from "Danmarks Miljøportal" (https://arealinformation.miljoeportal.dk/):

- p25_offentligareal.shp (§25 forests, accessed on on 5 April 2019
- skov_kortlaegning_2016_2018.shp (§15 forests, accessed 24 September 2019).

Number of forests: 9044 (§25 forests: 2906; and §15 forests: 6138).

### Untouched forests and "aftaler om natur"

The two other components of the high quality forest annotations are vector 
boundaries from the untouched forests (private and public), as well as areas 
with agreements on nature ("aftaler om natur"). The vector boundaries of these
areas were retrieved from "Miljøgis - Ansøgning om skovtilskud for private"  (https://miljoegis3.mim.dk/spatialmapsecure?profile=privatskovtilskud):

- tilsagn17_st_uroert_skov_privat_tilskud.shp (untouched forests, accessed on 6 July 2021)
- tilsagn18_st_uroert_skov.shp (untouched forests, accessed on 6 July 2021)
- tilsagn19_st_uroert_skov_privat_tilskud.shp (untouched forests, accessed on 6 July 2021)
- tilsagn20_st_uroert_skov_privat_tilskud.shp (untouched forests, accessed on 6 July 2021)
- aftale_natur_tinglyst.shp (agreements on nature, accessed on 6 July 2021)

Number of forests: 356 ("untouched": `r 15 + 20 + 37 + 46`; "aftaler om natur": 238).

---

## Low quality forests

Total number of low quality forest polygons: 10697.

### "Ikke" §25 forests

These forests are forests that were considered for being designated as §25 forests,
but did not meet the requirements (e.g., after completion of the field survey).
The vector geometries for these forests were shared with us by Bjarne Aabrandt 
Jensen (Miljøstyrelsen) in a personal communication on 19 November 2019.

- ikkeP25_skov.shp (personal communication, 19 November 2019)

Number of forests: 5848.

### NST plantations 

These forests are plantations owned by Denmark's environment agency "Naturstyrelsen" (NST).
The vector geometries and auxiliary data for these forests were obtained by personal 
communication from Bjørn Ole Ejlersen at NST to Pil Pedersen on 11 June 2020.

The source dataset includes all forests owned by NST. To subset only forests 
that are plantations, we filtered the data by excluding all forests that had 
an "ANV 4" value of 1, were classified as "urørt" or designated as "historical". 
We then sub-sampled the plantations to ensure a balanced training dataset between 
high and low quality forests (target :~10k high & ~10k low quality forests). We drew a sample of 5000 plantations. To account for variation 
within stand ages, we stratified the sample based on the following stand ages 
classes (years): [0, 10], (10, 25], (25, 50], (50, 75], (75, ∞). For each stand 
age we drew 1000 forests at random. Not all forests that were drawn in the sample
had an associated polygon in the separate vector geometry file (n missing = 151), 
these forests were not included in the final NST plantation subset.

- NST  2019 08012019 ber 16012020 til bios_au.xlsx (NST forests data table)
- LitraPolygoner_region.shp (NST forests polygons)

Number of forests: 4849.

---

## Cleaning and preparation of geometries

We observed some overlap between the assembled annotations for high and low
quality forests, both duplicate mappings within each category, but also some
duplicate mapping of parts of forests as high and low quality. These inaccuracies
were expected given the extent of the dataset and the fact that it was assembled
from multiple sources. To address the issue we carried out a systematic cleaning
of the geometries (fully reproducible through our source code). 

First, we removed internal overlap within the high quality geometries. For this
we iteratively removed all internal overlap starting with the p25 forests (internally sorted in arbitary order^), followed 
by the private old growth and lastly the p15 forests. We also buffered the geometries,
using a negative buffer of 10 cm to avoid line-overlaps due to geometric inaccuracies. Finally, we removed one
remaining p15 forest that failed to be filtered out. 

Second, we removed all internal overlap within the low quality geometries, taking
the same approach as for the high quality geometries except working fully in an arbitrary order. 
While doing that we also checked for overlap with any high quality geometries and 
if that was the case removed those. We also buffered the geometries,
using a negative buffer of 10 cm to avoid line-overlaps due to geometric inaccuracies.
Finally, we reomved one remaining ikke_p25 forest that was dublicated in the dataset. 

The resulting dataset consists of two type of forest annotations (8915 high and 9720 low quality polygons)
with no internal overlap within or between categories. 

---

## Pixel training dataset 

### Pixel sampling

We used the forest annotations (8915 high quality forests and 9720 low
quality forests) to generate a sample of 60k pixels based on the EcoDes-DK14 grid
to train our forest quality models. 

The EcoDes-DK15 dataset uses a version of the Danish national grid that divides 
terrestrial Denmark into 10 m x 10 m cells / pixels (UTM32). 
For the training dataset we drew a sample of 30k pixels each from within 
the high quality and the low quality forest polygons. Specifically, the sample 
was based on the EcoDes-DK15 "dtm10_m" descriptor raster). The sample was drawn
in the following fashion: First, we drew a random pixel from within each forest 
polygon (high or low), we then filled in the missing number of pixels to make up 
30k for each forest quality class (high or low). The filing step was done completely at
random, drawing from all remaining pixels available per class. The final dataset of pixels
therefore contained 30k unique pixels from each class. 

``` {r training_data_map, echo = FALSE, cache = TRUE}
# Load pixel sample
load("../data/training_data/pixel_training.Rda")

# Get Denmark boundaries
denmark <- ne_countries(scale = "large", 
                        country = "denmark",
                        returnclass = "sf") %>%
  st_transform(st_crs(pixel_training_data))

# Plot training data
ggplot() +
  geom_sf(data = denmark, fill = "NA", colour = "black") +
  geom_sf(data = pixel_training_data, 
          mapping = aes(colour = forest_value),
          size = 0.01) +
  labs(title = "Pixel sample for training") +
  coord_sf(datum = st_crs(pixel_training_data)) +
  theme_void()
```

Note, that because of the sampling scheme some pixels will be drawn from the 
individual forest polygons and the sampled pixels are therefore to some degree 
not statistically independent (as would samples from neighbouring polygons). However, the aim 
of our project is not to carry out a statistical analysis, but to train a 
machine learning classifier for predicting forest quality. The reduction in 
independence of the samples is therefore not an issue, as long as it does not lead
to over fitting of the models (which it did not). Instead sub-sampling of the 
forest polygons is actually desirable as it allows us to capture the variation 
within the forest polygons themselves and therefore increase the predictive 
capabilities of our models. We also include focal (window) predictors variables 
to account for within landscape-scale (100 m x 100 m) variation in our models. 

Finally, we chose a sample size of 60k pixels as a compromise between available
computing power and model output performance. There are approximately 56.3 million forest 
pixels in Denmark and the training sample therefore represents ~0.1% of the 
total forest area in the country. 

### Training / validation split (geographic stratifiaction)

To allow for an independent validation of the model performance, we split the
data 80/20 (training / validation) before carrying out the training. To account for potential
geographical covariation in the training data we used a geographic 
stratification when carrying out the split. This means that the split was not 
conducted at random on the whole dataset, but randomly within regions
(80/20 in each region). We used two different stratification schema of Denmark (see below). 

#### Biowide stratification

This stratification was developed for the BIOWIDE project (Brunbjerg et al. 2019).
The geometries are not publicly available and were kindly shared with us by 
Ane Brunbjerg (personal communication on 1 September 2021). Some further clean 
up of the geometries was required on our end. We had to make sure the boundaries of 
geometries were flush among neighbouring regions and that the coastlines were buffered. 

The stratification divides Denmark into six regions. 

``` {r biowide_regions, echo = FALSE, cache = TRUE}
biowide_regions <- read_sf("../data/stratification/biowide_georegions/biowide_zones.shp")
ggplot() +
  geom_sf(data = biowide_regions, aes(fill = region))
```

### Derek's stratification

The second stratification was developed by co-author Derek Corcoran based on 
climatic and other ecological covariates. 

It divides Denmark into three regions. 

```{r dereks_stratificaton, echo = FALSE, cache = TRUE}
dereks_strat <- rast("../data/stratification/derek_stratification/Results_2clim_5soil.tif") %>% as.polygons() %>% st_as_sf(crs = 25832) %>% mutate(region = factor(paste0("Region", Temp)))
ggplot() +
  geom_sf(data = dereks_strat, aes(fill = region))
```

## References

- Brunbjerg, A. K., Bruun, H. H., Brøndum, L., Classen, A. T., Dalby, L., Fog, K., Frøslev, T. G., Goldberg, I., Hansen, A. J., Hansen, M. D. D., Høye, T. T., Illum, A. A., Læssøe, T., Newman, G. S., Skipper, L., Søchting, U., & Ejrnæs, R. (2019). A systematic survey of regional multi-taxon biodiversity: Evaluating strategies and coverage. BMC Ecology, 19(1), 43. https://doi.org/10.1186/s12898-019-0260-x






