---
title: "DK Forest LiDAR - Forest Annotations & Training Data"
author: Jakob J. Assmann
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(sf)
library(tidyverse)
library(ggplot2)
library(terra)
library(rnaturalearth)
```

This document provides an overview of the forest annotations used for generating the 
training dataset that forms the base of our forest quality models for Denmark. 

These annotations are vector polygons of forests in Denmark that are of known 
"high" or "low" quality. We used these polygons to generate a training dataset 
of 30k pixels based on the 10 m grid of Denmark that is used by our models. 
The grid is defined by the EcoDes-DK15 dataset. A brief description on how the
final pixel training dataset was generated from the forest 
annotations can be found at the end of this document. 

Note: What makes a "high" or "low" quality forest is to some degree arbitrary. 
Our definitions here are the result of a long discussion and have been developed over
multiple years. The aim was to arrive at a workable definition that aligns 
with the current framework of forest designations in Denmark, while also ensuring that 
enough training data is available. We appreciate that our chosen definitions of 
forest quality are a simplification and not without flaws (e.g., we assume that 
all "plantations" are of low forest quality). 

---

## High quality forests

Total number of high quality forest polygons: 9400.

### §25 and §15 forest

The core of the high quality forest annotations is made up by the polygons for 
the designated §25 ("naturmaessigt saerlig vaerdifuld skov") and §15 ("skovnatur") forests. 
The vector boundaries of these forests were retrieved from "Danmarks Miljøportal" (https://arealinformation.miljoeportal.dk/):

- p25_offentligareal.shp (§25 forests, accessed on on 5 April 2019
- skov_kortlaegning_2016_2018.shp (§15 forests, accessed 24 September 2019).

Number of forests: 9044 (§25 forests: 2906; and §15 forests: 6138).

### Untouched forests and "aftaler om natur"

The two other components of the high quality forest annotations are vector 
boundaries from the untouched forests (private and public), as well as areas 
with agreements on nature ("aftaler om natur"). The vector boundaries of these
areas were retrieved from "Miljøgis - Ansøgning om skovtilskud for private"  (https://miljoegis3.mim.dk/spatialmapsecure?profile=privatskovtilskud):

- tilsagn17_st_uroert_skov_privat_tilskud.shp (untouched forests, accessed on 6 July 2021)
- tilsagn18_st_uroert_skov.shp (untouched forests, accessed on 6 July 2021)
- tilsagn19_st_uroert_skov_privat_tilskud.shp (untouched forests, accessed on 6 July 2021)
- tilsagn20_st_uroert_skov_privat_tilskud.shp (untouched forests, accessed on 6 July 2021)
- aftale_natur_tinglyst.shp (agreements on nature, accessed on 6 July 2021)

Number of forests: 356 ("untouched": `r 15 + 20 + 37 + 46`; "aftaler om natur": 238).

---

## Low quality forests

Total number of low quality forest polygons: 10697.

### "Ikke" §25 forests

These forests are forests that were considered for being designated as §25 forests,
but did not meet the requirements (e.g., after completion of the field survey).
The vector geometries for these forests were shared with us by Bjarne Aabrandt 
Jensen (Miljøstyrelsen) in a personal communication on 19 November 2019.

- ikkeP25_skov.shp (personal communication, 19 November 2019)

Number of forests: 5848.

### NST plantations 

These forests are plantations owned by Denmark's environment agency "Naturstyrelsen" (NST).
The vector geometries and auxiliary data for these forests were obtained by personal 
communication from Bjørn Ole Ejlersen at NST to Pil Pedersen on 11 June 2020.

The source dataset includes all forests owned by NST. To subset only forests 
that are plantations, we filtered the data by excluding all forests that had 
an "ANV 4" value of 1, were classified as "urørt" or designated as "historical". 
We then sub-sampled the plantations to ensure a balanced training dataset between 
high and low quality forests (target :~10k high & ~10k low quality forests). We drew a sample of 5000 plantations. To account for variation 
within stand ages, we stratified the sample based on the following stand ages 
classes (years): [0, 10], (10, 25], (25, 50], (50, 75], (75, ∞). For each stand 
age we drew 1000 forests at random. Not all forests that were drawn in the sample
had an associated polygon in the separate vector geometry file (n missing = 151), 
these forests were not included in the final NST plantation subset.

- NST  2019 08012019 ber 16012020 til bios_au.xlsx (NST forests data table)
- LitraPolygoner_region.shp (NST forests polygons)

Number of forests: 4849.

---

## Pixel training dataset 

### Pixel sampling

We used the forest annotations (9400 high quality forests and 10697 low
quality forests) to generate a sample of 200k pixels based on the EcoDes-DK14 grid
to train our forest quality models. 

The EcoDes-DK15 dataset uses a version of the Danish national grid that divides 
terrestrial Denmark into 10 m x 10 m cells / pixels (UTM32). 
For the training dataset we drew a sample of 100k pixels each from within 
the high quality and the low quality forest polygons. Specifically, the sample 
was based on the EcoDes-DK15 "dtm10_m" descriptor raster). The sample was drawn 
completely at random.

``` {r training_data_map, echo = FALSE, cache = TRUE}
# Load pixel sample
load("../data/training_data/pixel_training.Rda")

# Get Denmark boundaries
denmark <- ne_countries(scale = "large", 
                        country = "denmark",
                        returnclass = "sf") %>%
  st_transform(st_crs(pixel_training_data))

# Plot training data
ggplot() +
  geom_sf(data = denmark, fill = "NA", colour = "black") +
  geom_sf(data = pixel_training_data, 
          mapping = aes(colour = forest_value),
          size = 0.01) +
  labs(title = "Pixel sample for training") +
  coord_sf(datum = st_crs(pixel_training_data)) +
  theme_void()
```

Note, that because of the sampling scheme multiple pixels will be drawn from the 
individual forest polygons and the sampled pixels are therefore to some degree 
not statistically independent (as would samples from neighbouring polygons). However, the aim 
of our project is not to carry out a statistical analysis, but to train a 
machine learning classifier for predicting forest quality. The reduction in 
independence of the samples is therefore not an issue, as long as it does not lead
to over fitting of the models (which it did not). Instead sub-sampling of the 
forest polygons is actually desirable as it allows us to capture the variation 
within the forest polygons themselves and therefore increase the predictive 
capabilities of our models. We also include focal (window) predictors variables 
to account for within landscape-scale (100 m x 100 m) variation in our models. 

Finally, we chose a sample size of 200k pixels as a compromise between available
computing power and model performance. There are approximately 56.3 million forest 
pixels in Denmark and the training sample therefore represents ~0.36% of the 
total forest area in the country. 

### Training / validation split (geographic stratifiaction)

To allow for an independent validation of the model performance, we split the
data 80/20 (training / validation) before carrying out the training. To account for potential
geographical covariation in the training data we used a geographic 
stratification when carrying out the split. This means that the split was not 
conducted at random on the whole dataset, but randomly within regions
(80/20 in each region). We used two different stratification schema of Denmark (see below). 

#### Biowide stratification

This stratification was developed for the BIOWIDE project (Brunbjerg et al. 2019).
The geometries are not publicly available and were kindly shared with us by 
Ane Brunbjerg (personal communication on 1 September 2021). Some further clean 
up of the geometries was required on our end. We had to make sure the boundaries of 
geometries were flush among neighbouring regions and that the coastlines were buffered. 

The stratification divides Denmark into six regions. 

``` {r biowide_regions, echo = FALSE, cache = TRUE}
biowide_regions <- read_sf("../data/stratification/biowide_georegions/biowide_zones.shp")
ggplot() +
  geom_sf(data = biowide_regions, aes(fill = region))
```

### Derek's stratification

The second stratification was developed by co-author Derek Corcoran based on 
climatic and other ecological covariates. 

It divides Denmark into three regions. 

```{r dereks_stratificaton, echo = FALSE, cache = TRUE}
dereks_strat <- rast("../data/stratification/derek_stratification/Results_2clim_5soil.tif") %>% as.polygons() %>% st_as_sf(crs = 25832) %>% mutate(region = factor(paste0("Region", Temp)))
ggplot() +
  geom_sf(data = dereks_strat, aes(fill = region))
```

## References

- Brunbjerg, A. K., Bruun, H. H., Brøndum, L., Classen, A. T., Dalby, L., Fog, K., Frøslev, T. G., Goldberg, I., Hansen, A. J., Hansen, M. D. D., Høye, T. T., Illum, A. A., Læssøe, T., Newman, G. S., Skipper, L., Søchting, U., & Ejrnæs, R. (2019). A systematic survey of regional multi-taxon biodiversity: Evaluating strategies and coverage. BMC Ecology, 19(1), 43. https://doi.org/10.1186/s12898-019-0260-x






